---
title: "Patterns of Support in Minneapolis' Police Charter Amendment"
author: "Ryan Larson and Ryan LeCount"
date: "`r Sys.Date()`"
output:
  beamer_presentation:
    theme: Boadilla
    colortheme: beaver
    slide_level: 2
  ioslides_presentation: default
institute: |
  | Criminal Justice & Forensic Science
  | Social Justice & Social Change
  | Hamline University
urlcolor: blue
header-includes:
- \addtobeamertemplate{title page}{\centering \includegraphics[height=2cm, width=2cm]{hamline_logo.png}}{}
- \pgfdeclareimage[height=1cm, width=1.5cm]{piper}{piper.png}
- \logo{\pgfuseimage{piper}}
- \setbeamertemplate{itemize items}{>}
- \setbeamertemplate{itemize subitem}[>]
- \setbeamertemplate{itemize subsubitem}[>]
- \setbeamercolor{itemize item}{fg=red}
- \setbeamercolor{itemize subitem}{fg=red}
- \setbeamercolor{itemize subsubitem}{fg=red}
- \setbeamercolor{section name}{fg=white}
- \usepackage{dcolumn}
---

```{r message = FALSE, warning = FALSE, echo=FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
knitr::opts_chunk$set(warning=FALSE, message=FALSE, size = "tiny", out.width = "90%")
```

```{r setup, include=FALSE, message=F}
library(readr)
library(dplyr)
library(tidyr)
library(ggmap)
library(rvest)
library(stringr)
library(sf)
library(tigris)
library(tidycensus)
library(lubridate)
library(geomander)
```


# Introduction

## Minneapolis Charter Amendment 2


- amendment to Minneapolis city charter
  - replacement of MPD with Department of Public Safety
  - remove minimum police funding requirements and exclusive mayoral control
  - DPS commissioner controlled by mayor and city council
  - Failed: ~56% no
- Minneapolis is not the only municipality that had a police related ballot measures
  - Denver, Austin, Cleveland, Detroit, etc.
- we ask what drove support for CA2
  - political divides?
  - demographics (e.g., race)?
  - policing?
  - crime?

## Motivation: policing

- interpersonal contact between government officials a fundamental part of democratic political socialization (Soss 1999)
- police legitimacy
  - impacts perceptions of police violence (Jackson et a. 2013)
- legal cynicism (Sampson and Bartusch 1989)
  - deep-seated belief in incompetence, illegitimacy, and unresponsiveness of CJS (Kirk and Papachristos 2011)
  - grounded in negative encounters with LE and cultural processes (Carr et al. 2007; Brunson 2007)
  - police violence leads to lower 911 calls, particularly in Black communities (Desmond et al. 2016)
- over- and under-policing
  - simultaneous over- and under-policing in disadvantage MPLS neighborhoods (Phelps, Powell, and Robertson 2020)
  - race predicts higher agreements with both in PHDCN data (Boehme et al. 2020)

## Motivation: more than just policy

- political ideology
  - concern for biased policing linked to democratic voting (Drakulich et al. 2019)
    - Trump votes linked to police supporters with high levels of racial resentment 
  - GSS: conservative affiliation = more support (Roscigno and Preito-Hodge 2021)
- demographics
  - Black residents more negative attitudes about police, net of experiences (Wheelock et al. 2019)
  - race differences heightened amongst low SES individuals (Panditharatne et al. 2021)
  - residential stability linked to higher trust in police (Lee et al. 2019)
- practical concerns
  - perceptions of safety and past victimization predict attitudes toward police (Wheelock et al. 2019)

## Current Study

- conceptualize CA2 as not merely a policy issue
  - proxy for legal cynicism, etc.
  - experiences with police
  - political symbol and expression
  - concern with crime
  
- examine the voting patterns of CA2 along the lines of
  - demographics - race, median HH income, residential stability
  - indicators of policing
  - crime
  - political ideology

# Data and Methods

## Data: Voting-District Level

- Minnesota Secretary of State Voting Results
  - scraped CA2 results
  - scraped 2020 presidential election results
- Decennial Census 2020 P.L. 94-171 Redistricting Data
  - racial demographics
- American Community Survey 2020 5-Year Data
  - tract data converted to VD data using population-weighted interpolation
    - intersection between VD and CT, spatial join w/ blocks as points
    - VD measure estimated as a weighted sum of the block points of overlap by population size
  - used to gather economic and other demographic information
- Minneapolis Police Department Data - spatially located
  - uses of force, stops, officer-involved shootings

## Analytical Strategy

- spatial visualization of CA2 support
  - choropleth map of support by voting precinct
- scatterplots and bivariate correlation
  - focal measures of interest and support
- preliminary multivariate autoregressive model
  - $y = \rho_{lag}Wy+X\beta+\epsilon$
  - controls for spatial autocorrelation of CA2 support across "queen continguous" precincts
  - `spatialreg` package in `R`

```{r, include=FALSE}
census_20 <- get_decennial(
  geography = "voting district",
   variables = c("P4_001N", "P4_002N","P4_005N",
                 "P4_006N", "P4_007N", "P4_008N",
                 "P4_009N", "P4_010N", "P4_011N"),
  year = 2020,
  state = "MN",
  county = "Hennepin",
  geometry = TRUE,
  output = "wide"
 ) %>%
  rename(total_pop = P4_001N,
         hisp_pop = P4_002N,
         white_pop = P4_005N,
         black_pop = P4_006N,
         na_pop = P4_007N,
         asian_pop = P4_008N,
         nhpi_pop = P4_009N,
         other_pop = P4_010N,
         two_pop = P4_011N) %>%
  filter(str_detect(NAME, "Minneapolis")) %>%
  mutate(NAME = gsub("\\,.*", "", NAME),
         black_perc = black_pop/total_pop*100,
         hisp_perc = hisp_pop/total_pop*100,
         white_perc = white_pop/total_pop*100,
         na_perc = na_pop/total_pop*100,
         asian_perc = asian_pop/total_pop*100,
         other_perc = (nhpi_pop+other_pop+two_pop)/total_pop*100)
```

```{r, include=FALSE}
url_data <- "https://electionresults.sos.state.mn.us/Results/Index?ersElectionId=142&scenario=ResultsByPrecinctCrosstab&QuestionId=1323"

css_selector <- "#dblscroll > table"

precinct <- url_data %>% 
  read_html() %>% 
  html_element(css = css_selector) %>% 
  html_table() %>%
  slice(-n()) %>%
  rename(precinct = `County: Precinct`,
         yes_votes = `NP YES`,
         no_votes = `NP NO`) %>%
  mutate(precinct = str_remove(precinct, pattern = "Hennepin: "),
         yes_votes = str_remove(yes_votes, pattern = ","),
         no_votes = str_remove(no_votes, pattern = ",")) %>%
  mutate_at(c("yes_votes", "no_votes"), as.numeric) %>%
  mutate(perc_yes = yes_votes/(yes_votes+no_votes)*100) %>%
  mutate(precinct = str_replace_all(precinct, "MINNEAPOLIS", "Minneapolis"),
         precinct = str_replace_all(precinct, "-0", "-")) %>%
  mutate_all(~ifelse(is.nan(.), NA, .))

setequal(census_20$NAME, precinct$precinct) #same precincts in each


#join to census extract

precinct_sf <- census_20 %>%
  left_join(precinct, by = c("NAME"="precinct"))

#create north/south indicator?

#2020 presidential data

prez <- readxl::read_xlsx("MPLS Charter Amendment/Data/2020-general-federal-state-results-by-precinct-official.xlsx",
                          sheet = 2) %>%
  filter(MCDNAME=="Minneapolis") %>%
  select(PCTNAME, TOTVOTING, USPRSR, USPRSDFL) %>%
  mutate(pct_dem = ifelse(TOTVOTING!=0, USPRSDFL/TOTVOTING*100, NA_integer_),
         PCTNAME = str_replace_all(PCTNAME, "MINNEAPOLIS", "Minneapolis"),
         PCTNAME = str_replace_all(PCTNAME, "-01", "-1"),
         PCTNAME = str_replace_all(PCTNAME, "-02", "-2"),
         PCTNAME = str_replace_all(PCTNAME, "-03", "-3"),
         PCTNAME = str_replace_all(PCTNAME, "-04", "-4"),
         PCTNAME = str_replace_all(PCTNAME, "-05", "-5"),
         PCTNAME = str_replace_all(PCTNAME, "-06", "-6"),
         PCTNAME = str_replace_all(PCTNAME, "-07", "-7"),
         PCTNAME = str_replace_all(PCTNAME, "-08", "-8"),
         PCTNAME = str_replace_all(PCTNAME, "-09", "-9"))

precinct_sf <- precinct_sf %>%
  left_join(prez, by = c("NAME" = "PCTNAME"))

```

```{r, include=FALSE}
#Minneapolis Police Department - Use of Force Dashboard
uof_spatial <- read_csv("MPLS Charter Amendment/Data/Police_Use_of_Force.csv")  %>%
  mutate(date=ymd_hms(ResponseDate)) %>%
  filter(date >= as.Date("2020-01-01") & date < as.Date("2021-11-08")) %>%
  select(OBJECTID, X, Y, Race) %>%
  st_as_sf(coords = c("X", "Y"), crs = "NAD83", remove=F) %>%
  st_join(precinct_sf) %>%
  st_drop_geometry() %>%
  filter(!is.na(NAME)) %>%
  group_by(NAME, Race, .drop=F) %>%
  tally(name = "use_of_force") %>%
  filter(!is.na(Race)) %>%
  ungroup() %>%
  complete(NAME=precinct_sf$NAME, Race, fill = list(use_of_force = 0)) %>%
  arrange(NAME, Race) %>%
  mutate(race = str_to_lower(Race)) %>%
  select(-Race) %>%
  pivot_wider(names_from = race, 
              values_from = use_of_force,
              values_fill = 0,
              names_glue = "{race}_{.value}") %>%
  mutate(total_use_of_force = asian_use_of_force+black_use_of_force+`native american_use_of_force`+
           `not recorded_use_of_force`+`pacific islander_use_of_force`+unknown_use_of_force+           
            white_use_of_force)

# MPD Crime Data
crime_spatial <- read_csv("MPLS Charter Amendment/Data/Crime_Data.csv")  %>%
  mutate(date=ymd_hms(Occurred_Date)) %>%
  filter(date >= as.Date("2020-01-01") & date < as.Date("2021-11-08") &
           Type=="Crime Offenses (NIBRS)") %>%
  select(OBJECTID, Longitude, Latitude, Crime_Count) %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = "NAD83", remove=F) %>%
  st_join(precinct_sf) %>%
  st_drop_geometry() %>%
  filter(!is.na(NAME)) %>%
  group_by(NAME, .drop=F) %>%
  summarize(mpd_crime = sum(Crime_Count, na.rm = T)) %>%
  ungroup() %>%
  complete(NAME=precinct_sf$NAME, fill = list(mpd_crime = 0)) %>%
  arrange(NAME) 


#MPD Stop Dashboard
stops_spatial <- read_csv("MPLS Charter Amendment/Data/Police_Stop_Data.csv")  %>%
  mutate(date=ymd_hms(responseDate)) %>%
  filter(date >= as.Date("2020-01-01") & date < as.Date("2021-11-08")) %>%
  select(OBJECTID, long, lat, race) %>%
  st_as_sf(coords = c("long", "lat"), crs = "NAD83", remove=F) %>%
  st_join(precinct_sf) %>%
  st_drop_geometry() %>%
  filter(!is.na(NAME)) %>%
  group_by(NAME, race, .drop=F) %>%
  tally(name = "stops") %>%
  filter(!is.na(race)) %>%
  ungroup() %>%
  complete(NAME=precinct_sf$NAME, race, fill = list(stops = 0)) %>%
  arrange(NAME, race) %>%
  mutate(race = str_to_lower(race)) %>%
  pivot_wider(names_from = race, 
              values_from = stops,
              values_fill = 0,
              names_glue = "{race}_{.value}") %>%
  mutate(total_stops = asian_stops+black_stops+`native american_stops`+
           `east african_stops`+`latino_stops`+unknown_stops+other_stops+           
            white_stops)


#Officer Involved Shootings - MPD
ois_spatial <- read_csv("MPLS Charter Amendment/Data/Police_Officer_Involved_Shootings.csv") %>%
  mutate(date=ymd_hms(IncidentDate)) %>%
  filter(date >= as.Date("2020-01-01") & date < as.Date("2021-11-08")) %>%
  select(OBJECTID, CenterLatitude, CenterLongitude) %>%
  rename(lat = CenterLatitude,
         long = CenterLongitude) %>%
  st_as_sf(coords = c("long", "lat"), crs = "NAD83", remove=F) %>%
  st_join(precinct_sf) %>%
  st_drop_geometry() %>%
  filter(!is.na(NAME)) %>%
  group_by(NAME, .drop=F) %>%
  tally(name = "ois") %>%
  ungroup() %>%
  complete(NAME=precinct_sf$NAME, fill = list(ois = 0)) %>%
  arrange(NAME) 

precinct_sf <- precinct_sf %>%
  left_join(uof_spatial, by = "NAME") %>%
  left_join(stops_spatial, by = "NAME") %>%
  left_join(ois_spatial, by = "NAME") %>%
  left_join(crime_spatial, by = "NAME") %>%
  mutate(uof_rate = total_use_of_force/total_pop*1000,
         blk_uof_rate = black_use_of_force/black_pop*1000,
         blk_uof_prop = black_use_of_force/total_use_of_force,
         stops_rate = total_stops/total_pop*1000,
         blk_stop_rate = black_stops/black_pop*1000,
         blk_stop_prop = black_stops/total_stops,
         ois_rate = ois/total_pop*1000,
         crime_rate = mpd_crime/total_pop*1000)
```

```{r, include=FALSE}
#MN ACS 2020 covariates
mn_tract <- get_acs(geography = "tract", 
            state = "MN",
               variables = c("B01001_001E", "B03002_003E", "B03002_004E", "B03002_005E",
                             "B03002_006E", "B03002_007E", "B03002_008E", "B03002_009E",
                             "B03002_010E", "B03002_011E", "B03002_012E", "B23025_002E",
                             "B23025_005E", "B17001_002E", "B19057_002E", "B11003_015E", 
                             "B06009_002E", "B06009_005E", "C24010_001E", "C24010_003E",
                             "C24010_039E", "B11001_003E", "B01001_002E", "B05001_006E",
                             "B01001_003E", "B01001_004E", "B01001_005E", "B01001_006E",
                             "B01001_007E", "B01001_008E", "B01001_009E", "B01001_010E",
                             "B01001_011E", "B01001_012E", "B01001_013E", "B01001_014E",
                             "B01001_015E", "B01001_016E", "B01001_017E", "B01001_018E",
                             "B01001_019E", "B01001_020E", "B01001_021E", "B01001_022E",
                             "B01001_023E", "B01001_024E", "B01001_025E", "B01001_027E",
                             "B01001_028E", "B01001_029E", "B01001_030E", "B01001_031E",
                             "B01001_032E", "B01001_033E", "B01001_034E", "B01001_035E",
                             "B01001_036E", "B01001_037E", "B01001_038E", "B01001_039E",
                             "B01001_040E", "B01001_041E", "B01001_042E", "B01001_043E",
                             "B01001_044E", "B01001_045E", "B01001_046E", "B01001_047E", 
                             "B01001_048E", "B01001_049E", "B07001_017E", "B25003_002E",
                             "B05002_013E", "B19013_001E"),
               output = "wide",
               survey = "acs5",
               year = 2020,
            geometry = T) %>%
  select(-ends_with("M", ignore.case = F)) %>%
  rename(total_pop = B01001_001E, white_pop = B03002_003E, black_pop = B03002_004E,
         na_pop = B03002_005E, asian_pop = B03002_006E, hpi_pop = B03002_007E,
         other_pop = B03002_008E, biracial_pop = B03002_009E, biracial_other_pop = B03002_010E,
         biracial_three_pop = B03002_011E, hisp_pop = B03002_012E, total_ilf = B23025_002E,
         unemp = B23025_005E, povlevel = B17001_002E, pub_assist = B19057_002E,
         female_hh = B11003_015E, no_hs_dip = B06009_002E, bach_degree = B06009_005E,
         total_employed = C24010_001E, employed_mbsa_male = C24010_003E,
         employed_mbsa_female = C24010_039E, mar_fam = B11001_003E, male = B01001_002E,
         noncitizen = B05001_006E,
        age_m_5_under = B01001_003E, age_m_5_9 = B01001_004E, age_m_10_14 = B01001_005E, 
        age_m_15_17 = B01001_006E, age_m_18_19 = B01001_007E, age_m_20 = B01001_008E,
        age_m_21 = B01001_009E, age_m_22_24 = B01001_010E, age_m_25_29 = B01001_011E, 
        age_m_30_34 = B01001_012E, age_m_35_39 = B01001_013E, age_m_40_44 = B01001_014E,
        age_m_45_49 = B01001_015E, age_m_50_54 = B01001_016E, age_m_55_59 = B01001_017E, 
        age_m_60_61 = B01001_018E, age_m_62_64 = B01001_019E, age_m_65_66 = B01001_020E,
        age_m_67_69 = B01001_021E, age_m_70_74 = B01001_022E, age_m_75_79 = B01001_023E, 
        age_m_80_84 = B01001_024E, age_m_85_plus = B01001_025E, age_f_5_under = B01001_027E,
        age_f_5_9 = B01001_028E, age_f_10_14 = B01001_029E, age_f_15_17 = B01001_030E, 
        age_f_18_19 = B01001_031E, age_f_20 = B01001_032E, age_f_21 = B01001_033E,
        age_f_22_24 = B01001_034E, age_f_25_29 = B01001_035E, age_f_30_34 = B01001_036E, 
        age_f_35_39 = B01001_037E, age_f_40_44 = B01001_038E, age_f_45_49 = B01001_039E,
        age_f_50_54 = B01001_040E, age_f_55_59 = B01001_041E, age_f_60_61 = B01001_042E, 
        age_f_62_64 = B01001_043E, age_f_65_66 = B01001_044E, age_f_67_69 = B01001_045E, 
        age_f_70_74 = B01001_046E, age_f_75_79 = B01001_047E, age_f_80_84 = B01001_048E, 
        age_f_85_plus = B01001_049E, res_mob = B07001_017E, 
        own_hh = B25003_002E, foreign = B05002_013E,
        med_hh_inc = B19013_001E) %>%
  select(-c(white_pop, black_pop,
         na_pop, asian_pop, hpi_pop, other_pop, biracial_pop, biracial_other_pop,
         biracial_three_pop, hisp_pop))

#Minneapolis Shapefile
mpls <- st_read("MPLS Charter Amendment/Data/City_Boundary/City_Boundary.shp") %>%
   st_transform(st_crs(mn_tract))

mpls_tract <- mn_tract %>%
  st_filter(mpls, .predicate = st_intersects) %>%
    mutate(GEOID = as.numeric(GEOID),
         tract_area = as.numeric(st_area(.)),
         tract_area_sqkm = tract_area*.000001, 
         tract_area_sqmi = tract_area_sqkm*.386102,
         intersection_area = as.numeric(st_area(st_intersection(., mpls))),
         perc_intersection = intersection_area/tract_area*100) %>%
  filter(perc_intersection >= 2)

library(tigris)

hennepin_blocks <- blocks(
  "MN",
  "Hennepin",
  year = 2020
)

#VD ACS interpolation
acs_interp <- interpolate_pw(
  from = mpls_tract,
  to = precinct_sf,
  to_id = "GEOID",
  weights = hennepin_blocks,
  weight_column = "POP20",
  crs = 26993,
  extensive = TRUE) %>%
  mutate(unemp_rate = 100*unemp/total_ilf, 
         pov_rate = 100*povlevel/total_pop,
         pub_assist_rate = 100*pub_assist/total_pop,
         female_hh_rate = 100*female_hh/total_pop,
         no_hs_dip_rate = 100*no_hs_dip/total_pop,
         bach_degree_rate = 100*bach_degree/total_pop,
         employed_mbsa = employed_mbsa_male+employed_mbsa_female,
         employed_mbsa_rate = 100*employed_mbsa/total_employed,
         mar_fam_rate = 100*mar_fam/total_pop, 
         male_rate = 100*male/total_pop,
         noncitizen_rate = 100*noncitizen/total_pop,
         age_below_18_perc = 100*(age_m_5_under+age_f_5_under+age_m_5_9+
                                  age_f_5_9+age_m_10_14+age_f_10_14+age_m_15_17+
                                  age_f_15_17)/total_pop,
         age_19_29_perc = 100*(age_m_18_19+age_f_18_19+age_m_20+age_f_20+age_m_21+age_f_21+
                          age_m_22_24+age_f_22_24+age_m_25_29+age_f_25_29)/total_pop,
         age_30_49_perc = 100*(age_m_30_34+age_f_30_34+age_m_35_39+age_f_35_39+
                               age_m_40_44+age_f_40_44+age_m_45_49+age_f_45_49)/total_pop,
         age_50_69_perc = 100*(age_m_50_54+age_f_50_54+age_m_55_59+age_f_55_59+
                               age_m_60_61+age_f_60_61+age_m_62_64+age_f_62_64+
                               age_m_65_66+age_f_65_66+age_m_67_69+age_f_67_69)/total_pop,
         age_70_plus_perc = 100*(age_m_70_74+age_f_70_74+age_m_75_79+age_f_75_79+
                                 age_m_80_84+age_f_80_84+age_m_85_plus+age_f_85_plus)/total_pop,
         res_mob_rate = 100-100*res_mob/total_pop,
         own_hh_rate = 100*own_hh/total_pop,
         foreign_rate = 100*foreign/total_pop) %>%
  st_drop_geometry()

#join ACS data

precinct_sf <- precinct_sf %>%
  left_join(acs_interp, by = "GEOID")

#matches <- geo_match(from = precinct_sf, to = mpls_tract, method = 'centroid')
#est_med_hh_inc <- estimate_down(wts = precinct_sf$total_pop, value = mpls_tract$med_hh_inc, group = matches)
#precinct_sf$medhhinc_est <- est_med_hh_inc

```

# Results

## Spatial Patterns of Support

```{r, echo=FALSE}
ggplot() +
  geom_sf(data = precinct_sf, aes(geometry = geometry, fill = perc_yes)) + 
  scale_fill_distiller(palette = "RdYlGn")+
  labs(title = "Figure 1: Support for MPLS Charter Amendment #2, 2021", 
       subtitle = "MN Secretary of State",
       fill = "Percent Support")+
  theme_void()
```

## Bivariate Relationships: Race

```{r, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(precinct_sf, aes(y = perc_yes, x = black_perc))+
  geom_point()+
  geom_smooth(method = "loess", aes(color = "LOESS"))+
  geom_smooth(method = "lm", aes(color = "OLS"))+
  labs(y = "CA2 Support", x = "Percent Black",
       title = "Scatterplot of CA2 Support and Percent Black",
       color = "")+
  theme_classic()
```

## Bivariate Relationships: Class

```{r, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(precinct_sf, aes(y = perc_yes, x = med_hh_inc))+
  geom_point()+
  geom_smooth(method = "loess", aes(color = "LOESS"))+
  geom_smooth(method = "lm", aes(color = "OLS"))+
  labs(y = "CA2 Support", x = "Median HH Income",
       title = "Scatterplot of CA2 Support and Median HH Income",
       color = "")+
  theme_classic()
```

## Bivariate Relationships: Residential Stability

```{r, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(precinct_sf, aes(y = perc_yes, x = own_hh_rate))+
  geom_point()+
  geom_smooth(method = "loess", aes(color = "LOESS"))+
  geom_smooth(method = "lm", aes(color = "OLS"))+
  labs(y = "CA2 Support", x = "Percent Own HH",
       title = "Scatterplot of CA2 Support and HH Owndership Rate",
       color = "")+
  theme_classic()
```

## Bivariate Relationships: Political Ideology

```{r, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(precinct_sf, aes(y = perc_yes, x = pct_dem))+
  geom_point()+
  geom_smooth(method = "loess", aes(color = "LOESS"))+
  geom_smooth(method = "lm", aes(color = "OLS"))+
  labs(y = "CA2 Support", x = "Pct. Biden",
       title = "Scatterplot of CA2 Support and Percent Biden",
       color = "")+
  theme_classic()
```


## Bivariate Relationships: Uses of Force

```{r, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(precinct_sf, aes(y = perc_yes, x = uof_rate))+
  geom_point()+
  geom_smooth(method = "loess", aes(color = "LOESS"))+
  geom_smooth(method = "lm", aes(color = "OLS"))+
  labs(y = "CA2 Support", x = "UOF Rate",
       title = "Scatterplot of CA2 Support and UOF Rate",
       color = "")+
  theme_classic()
```

## Bivariate Relationships: Crime


```{r, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(precinct_sf, aes(y = perc_yes, x = crime_rate))+
  geom_point()+
  geom_smooth(method = "loess", aes(color = "LOESS"))+
  geom_smooth(method = "lm", aes(color = "OLS"))+
  labs(y = "CA2 Support", x = "Crime Rate",
       title = "Scatterplot of CA2 Support and Crime Rate",
       color = "")+
  theme_classic()
```


## Preliminary Multivariate Model

```{r, include = FALSE}
library(spdep)
library(spatialreg)
library(stargazer)

#neighbors: queen adjacency
mpls_nb <- poly2nb(precinct_sf, queen=T)
summary(mpls_nb)

#neighbor weights
mpls_w <- nb2listw(mpls_nb, style = "W", zero.policy = T)

#print(lm.morantest(m1, mpls_w)) #testing for spatial autocorrelation in residuals

m1_lag <- lagsarlm(perc_yes~black_perc+I(black_perc^2)+
                     log(med_hh_inc)*pct_dem+
                     uof_rate+crime_rate+
                     own_hh_rate,
                   data = precinct_sf, listw = mpls_w)
summary(m1_lag)
```

```{r, results='asis', warning=FALSE, echo = FALSE}
stargazer(m1_lag,
           title = "Spatial AR(1) Model of MPLS CA2 Support",
          covariate.labels = c("Percent Black","Percent Black Sq. ",
                               "Percent Biden",
                               "Median HH Income (log)",
                               "UOF Rate", "Crime Rate",  
                               "Own HH Rate", "Med. HH Inc. X Pct. Biden"),
          add.lines = list(c("SAR(1)", ".705^{***} (.061)")),
          dep.var.caption = "Percent Support",
          dep.var.labels.include = FALSE,
          model.numbers = FALSE,
          single.row = TRUE,
          align = T,
          header = FALSE,
          #omit.stat = "adj.rsq",
          star.cutoffs = c(.05, .01, .001), star.char = c("*","**","***"))
```

## Conclusions

- nonlinear patterns in black support
  - taper in support not just due to concerns with crime, police UOF
  - perhaps broader context of legal cynicism
- politicized issue
  - however, more advantaged individuals saw through a more ideological lens
  - the privilege to see this as political, as opposed to practical
- practical concerns of crime and police violence did not explain support
- future research
  - delve deeper into demographically relationships: interactions
  - race specific indicators of police violence
  - relative voter turnout as indicator of legal cynicism

## Thank you!

- Contact
  - Ryan Larson, Hamline University
  - rlarson21@hamline.edu
  - [@ryanplarson](https://twitter.com/ryanplarson)


  


