---
title: "Car Jacking MPLS - Tract"
author: "Ryan Larson"
date: "4/8/2022"
output: pdf_document
---



```{r setup, include=FALSE, message=F}
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(tidycensus)
library(sf)
library(purrr)
library(ggmap)
library(stargazer)
library(stringr)
library(lubridate)
library(ISOweek)
library(zoo)

#TO DO!
#old
#descriptive paper: over-time; tracts over time
#build dataset at the incident level; spatially join tract information 
#collective efficacy: any purchase on this? 
#time-varying indicators of pandemic, school closures, institution closures, social safety net, pandemic checks
#analysis of discrepancies between incidents and charges
#compare to homicides
#media stuff - coverage is responsible to rates in which places


# Research Q: what tracts experienced what kind of increase?
#is the rate/count above a certain threshold - any vs. none
#DV: none-to-any; multinomial?

#Dv: media attention - following raw change vs. different types of change in certain places

#t-tests of variables between strib coverage and no coverage
#build # media mentions outcome variables
#time-series weekly model predicting media mentions, counts in advantaged vs. disadvantaged places
  #multinomial predictors (0-1, 1-few, few-many?)
#(months/quarters for the plot) 



google_key <- readRDS("C:/Users/DELL/Documents/R/Rworkingdirectory/google_api_key.rds")
register_google(key = google_key)
```

# Spatial Data

```{r, message=FALSE, warning=FALSE}
#MN tracts
tracts <- get_acs(geography = "tract", 
                  state = "MN",
                variables = "B01001_001E",
                output = "wide",
                survey = "acs5",
                year = 2020,
                geometry = T)

#Minneapolis Shapefile
mpls <- st_read("Data/mpls_city-shp/16cdbbfa-ad10-493c-afaf-52b61f2e76e42020329-1-180h9ap.whbo.shp") %>%
   st_transform(st_crs(tracts))

mpls_tract <- tracts %>%
  st_filter(mpls, .predicate = st_intersects) %>%
  mutate(GEOID = as.numeric(GEOID),
         tract_area = as.numeric(st_area(.)),
         tract_area_sqkm = tract_area*.000001, 
         tract_area_sqmi = tract_area_sqkm*.386102,
         intersection_area = as.numeric(st_area(st_intersection(., mpls))),
         perc_intersection = intersection_area/tract_area*100) %>%
  filter(perc_intersection >= 2) %>%
  select(-"B01001_001M")

```

# ACS Covariates

```{r}
acs_17 <- get_acs(
  geography = "tract",
  variables = c("B01001_001E"),
  year = 2017,
  state = "MN",
  county = "Hennepin",
  geometry = TRUE) %>% 
  select(estimate)

acs_18 <- get_acs(
  geography = "tract",
  variables = c("B01001_001E"),
  year = 2018,
  state = "MN",
  county = "Hennepin",
  geometry = TRUE
) %>%
select(estimate)

acs_19 <- get_acs(
  geography = "tract",
  variables = c("B01001_001E"),
  year = 2019,
  state = "MN",
  county = "Hennepin",
  geometry = TRUE
) %>%
select(estimate)


acs_20 <- get_acs(
  geography = "tract",
  variables = c("B01001_001E"),
  year = 2020,
  state = "MN",
  county = "Hennepin",
  geometry = TRUE
 )

library(tigris)

hennepin_blocks <- blocks(
  "MN",
  "Hennepin",
  year = 2020
)

#2017 ACS interpolation
acs_1720 <- interpolate_pw(
  from = acs_17,
  to = acs_20,
  to_id = "GEOID",
  weights = hennepin_blocks,
  weight_column = "POP20",
  crs = 26993,
  extensive = TRUE) %>%
  mutate(year = 2017) 

#2018 ACS interpolation
acs_1820 <- interpolate_pw(
  from = acs_18,
  to = acs_20,
  to_id = "GEOID",
  weights = hennepin_blocks,
  weight_column = "POP20",
  crs = 26993,
  extensive = TRUE) %>%
  mutate(year = 2018)


#2019 ACS interpolation
acs_1920 <- interpolate_pw(
  from = acs_19,
  to = acs_20,
  to_id = "GEOID",
  weights = hennepin_blocks,
  weight_column = "POP20",
  crs = 26993,
  extensive = TRUE) %>%
  mutate(year = 2019)


pop_denoms <- acs_20 %>%
  st_transform(crs = 26993) %>%
  mutate(year = 2020) %>%
  select(-moe, -variable, -NAME) %>%
  rbind(acs_1720, acs_1820, acs_1920) %>%
  filter(GEOID %in% mpls_tract$GEOID)

#2021+2022: LOCF
pop_denom_21 <- pop_denoms %>%
  filter(year==2020) %>%
  select(GEOID, year, estimate) %>%
  mutate(year = 2021) 

pop_denom_22 <- pop_denoms %>%
  filter(year==2020) %>%
  select(GEOID, year, estimate) %>%
  mutate(year = 2022) 

pop_denom_locf <- pop_denoms %>% 
  rbind(pop_denom_21, pop_denom_22) %>%
  rename(total_pop = estimate) %>%
  mutate(GEOID = as.numeric(GEOID)) %>%
  st_drop_geometry()

#ACS 2020 L-2 covariates
acs_2020 <- get_acs(geography = "tract", 
            state = "MN",
               variables = c("B01001_001E", "B03002_003E", "B03002_004E", "B03002_005E",
                             "B03002_006E", "B03002_007E", "B03002_008E", "B03002_009E",
                             "B03002_010E", "B03002_011E", "B03002_012E", "B23025_002E",
                             "B23025_005E", "B17001_002E", "B19057_002E", "B11003_015E", 
                             "B06009_002E", "B06009_005E", "C24010_001E", "C24010_003E",
                             "C24010_039E", "B11001_003E", "B01001_002E", "B05001_006E",
                             "B01001_003E", "B01001_004E", "B01001_005E", "B01001_006E",
                             "B01001_007E", "B01001_008E", "B01001_009E", "B01001_010E",
                             "B01001_011E", "B01001_012E", "B01001_013E", "B01001_014E",
                             "B01001_015E", "B01001_016E", "B01001_017E", "B01001_018E",
                             "B01001_019E", "B01001_020E", "B01001_021E", "B01001_022E",
                             "B01001_023E", "B01001_024E", "B01001_025E", "B01001_027E",
                             "B01001_028E", "B01001_029E", "B01001_030E", "B01001_031E",
                             "B01001_032E", "B01001_033E", "B01001_034E", "B01001_035E",
                             "B01001_036E", "B01001_037E", "B01001_038E", "B01001_039E",
                             "B01001_040E", "B01001_041E", "B01001_042E", "B01001_043E",
                             "B01001_044E", "B01001_045E", "B01001_046E", "B01001_047E", 
                             "B01001_048E", "B01001_049E", "B07001_017E", "B25003_002E",
                             "B05002_013E", "B19013_001E"),
               output = "wide",
               survey = "acs5",
               year = 2020) %>%
  select(-ends_with("M", ignore.case = F)) %>%
  rename(total_pop = B01001_001E, white_pop = B03002_003E, black_pop = B03002_004E,
         na_pop = B03002_005E, asian_pop = B03002_006E, hpi_pop = B03002_007E,
         other_pop = B03002_008E, biracial_pop = B03002_009E, biracial_other_pop = B03002_010E,
         biracial_three_pop = B03002_011E, hisp_pop = B03002_012E, total_ilf = B23025_002E,
         unemp = B23025_005E, povlevel = B17001_002E, pub_assist = B19057_002E,
         female_hh = B11003_015E, no_hs_dip = B06009_002E, bach_degree = B06009_005E,
         total_employed = C24010_001E, employed_mbsa_male = C24010_003E,
         employed_mbsa_female = C24010_039E, mar_fam = B11001_003E, male = B01001_002E,
         noncitizen = B05001_006E,
        age_m_5_under = B01001_003E, age_m_5_9 = B01001_004E, age_m_10_14 = B01001_005E, 
        age_m_15_17 = B01001_006E, age_m_18_19 = B01001_007E, age_m_20 = B01001_008E,
        age_m_21 = B01001_009E, age_m_22_24 = B01001_010E, age_m_25_29 = B01001_011E, 
        age_m_30_34 = B01001_012E, age_m_35_39 = B01001_013E, age_m_40_44 = B01001_014E,
        age_m_45_49 = B01001_015E, age_m_50_54 = B01001_016E, age_m_55_59 = B01001_017E, 
        age_m_60_61 = B01001_018E, age_m_62_64 = B01001_019E, age_m_65_66 = B01001_020E,
        age_m_67_69 = B01001_021E, age_m_70_74 = B01001_022E, age_m_75_79 = B01001_023E, 
        age_m_80_84 = B01001_024E, age_m_85_plus = B01001_025E, age_f_5_under = B01001_027E,
        age_f_5_9 = B01001_028E, age_f_10_14 = B01001_029E, age_f_15_17 = B01001_030E, 
        age_f_18_19 = B01001_031E, age_f_20 = B01001_032E, age_f_21 = B01001_033E,
        age_f_22_24 = B01001_034E, age_f_25_29 = B01001_035E, age_f_30_34 = B01001_036E, 
        age_f_35_39 = B01001_037E, age_f_40_44 = B01001_038E, age_f_45_49 = B01001_039E,
        age_f_50_54 = B01001_040E, age_f_55_59 = B01001_041E, age_f_60_61 = B01001_042E, 
        age_f_62_64 = B01001_043E, age_f_65_66 = B01001_044E, age_f_67_69 = B01001_045E, 
        age_f_70_74 = B01001_046E, age_f_75_79 = B01001_047E, age_f_80_84 = B01001_048E, 
        age_f_85_plus = B01001_049E, res_mob = B07001_017E, 
        own_hh = B25003_002E, foreign = B05002_013E,
        med_hh_inc = B19013_001E) %>%
  mutate(white_prop = white_pop/total_pop,
         black_prop = black_pop/total_pop,
         na_prop = na_pop/total_pop,
         asian_prop = asian_pop/total_pop,
         hpi_prop = hpi_pop/total_pop,
         other_prop = other_pop/total_pop,
         biracial_prop = (biracial_pop+biracial_other_pop+biracial_three_pop)/total_pop,
         hisp_prop = hisp_pop/total_pop,
         white_perc = 100*white_pop/total_pop,
         black_perc = 100*black_pop/total_pop,
         na_perc = 100*na_pop/total_pop,
         asian_perc = 100*asian_pop/total_pop,
         hpi_perc = 100*hpi_pop/total_pop,
         other_perc = 100*other_pop/total_pop,
         biracial_perc = 100*(biracial_pop+biracial_other_pop+biracial_three_pop)/total_pop,
         hisp_perc = 100*hisp_pop/total_pop,
         unemp_rate = 100*unemp/total_ilf, 
         pov_rate = 100*povlevel/total_pop,
         pub_assist_rate = 100*pub_assist/total_pop,
         female_hh_rate = 100*female_hh/total_pop,
         no_hs_dip_rate = 100*no_hs_dip/total_pop,
         bach_degree_rate = 100*bach_degree/total_pop,
         employed_mbsa = employed_mbsa_male+employed_mbsa_female,
         employed_mbsa_rate = 100*employed_mbsa/total_employed,
         mar_fam_rate = 100*mar_fam/total_pop, 
         male_rate = 100*male/total_pop,
         noncitizen_rate = 100*noncitizen/total_pop,
         race_eth_hetero  = 1-(white_prop^2+black_prop^2+na_prop^2+asian_prop^2+
                       hpi_prop^2+other_prop^2+other_prop^2+biracial_prop^2+hisp_prop^2),
         age_below_18_perc = 100*(age_m_5_under+age_f_5_under+age_m_5_9+
                                  age_f_5_9+age_m_10_14+age_f_10_14+age_m_15_17+
                                  age_f_15_17)/total_pop,
         age_19_29_perc = 100*(age_m_18_19+age_f_18_19+age_m_20+age_f_20+age_m_21+age_f_21+
                          age_m_22_24+age_f_22_24+age_m_25_29+age_f_25_29)/total_pop,
         age_30_49_perc = 100*(age_m_30_34+age_f_30_34+age_m_35_39+age_f_35_39+
                               age_m_40_44+age_f_40_44+age_m_45_49+age_f_45_49)/total_pop,
         age_50_69_perc = 100*(age_m_50_54+age_f_50_54+age_m_55_59+age_f_55_59+
                               age_m_60_61+age_f_60_61+age_m_62_64+age_f_62_64+
                               age_m_65_66+age_f_65_66+age_m_67_69+age_f_67_69)/total_pop,
         age_70_plus_perc = 100*(age_m_70_74+age_f_70_74+age_m_75_79+age_f_75_79+
                                 age_m_80_84+age_f_80_84+age_m_85_plus+age_f_85_plus)/total_pop,
         res_mob_rate = 100-100*res_mob/total_pop,
         own_hh_rate = 100*own_hh/total_pop,
         foreign_rate = 100*foreign/total_pop) 
```

# Open Minneapolis Carjacking Data

```{r, message=FALSE}
#open minneapolis crime data 2019-4/7 (date of download)
cj_spatial <- read_csv("Car Jacking/crime_data.csv")  %>%
  filter(Offense=="Carjacking - Subset of Robbery") %>% #filter carjackings
  mutate(date=ymd_hms(Occurred_Date),
         year=isoyear(date),
         week=isoweek(date)) %>%
  select(OBJECTID, year, week, Latitude, Longitude) %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = "NAD83", remove=F) %>%
  st_join(mpls_tract) %>% #spatial join neighborhoods
  st_drop_geometry() %>%
  filter(!is.na(GEOID)) %>%
  group_by(year, week, GEOID, .drop=F) %>%
  tally(name = "car_jack") %>%
  ungroup() %>%
  complete(year, week, GEOID=mpls_tract$GEOID, fill = list(car_jack = 0)) %>%
  filter(!(year==2022 & week >= 14) & #removing unobserved/redundant completions
           !(year==2021 & week==53) & #removing unobserved/redundant completions
           !(year==2020 & week < 36)) %>% #removing weeks before Sept. 2020, isoweek 36
  arrange(GEOID, year, week) %>%
  left_join(mpls_tract, by = "GEOID") %>%
  left_join(pop_denom_locf, by = c("GEOID", "year")) %>%
  mutate(car_jack_rate = car_jack/total_pop*1000) %>%
  st_as_sf()
```


# MPLS Carjackings by Week

```{r}
#aggregate to week over tracts
cj_week <- cj_spatial %>%
  group_by(year, week) %>%
  summarize(car_jack = sum(car_jack, na.rm = T),
            total_pop = sum(total_pop, na.rm = T)) %>%
  mutate(begin_date = ISOweek2date(paste(year, paste0("W", sprintf("%02d", week)), 1,sep = "-")),
         end_date = begin_date+weeks(1)-days(1),
         car_jack_rate = car_jack/total_pop*1000) 
  
ggplot(cj_week)+
  geom_line(aes(x=begin_date, y=car_jack_rate))+ 
  scale_x_date(date_labels = "%b-%Y", 
               limits = c(min(cj_week$begin_date), max(cj_week$begin_date)))+
  labs(title = "Figure X: Weekly Minneapolis Carjackings, 8/31/2020-4/7/2022",
       subtitle = "openminneapolis Crime Data",
       x = "Week",
       y = "Weekly Carjacking Rate/ 1,000")+
  theme_minimal()+
    theme(axis.text.x=element_text(angle=45, hjust=1)) 

```


# MPLS ZCTA Carjackings Map

```{r}
#aggregate to neighborhood-year level
cj_tract_year <- cj_spatial %>%
  group_by(GEOID, year) %>%
  summarize(car_jack = sum(car_jack, na.rm = T),
            total_pop = sum(B01001_001E, na.rm = T),
            car_jack_rate = car_jack/total_pop*1000) 


ggplot() +
  geom_sf(data = cj_tract_year, aes(geometry = geometry, fill = car_jack_rate)) + 
  geom_sf(data = mpls, aes(geometry = geometry), color = "black", alpha = 0)+
  facet_wrap(~year)+
  scale_fill_distiller(palette = "Spectral")+
  labs(title = "Figure X: Minneapolis Carjacking Rates by Tract and Year", 
       subtitle = "openminneapolis Crime Data",
       fill = "Weekly Rate/ 1,000")+
  theme_void()
```



# Expanded MPLS Carjacking (Crime Incidents) Data

```{r}
cj_exp <- read_csv("Car Jacking/MPDdata_051022.csv") %>%
   mutate(date=mdy_hm(reporteddate),
         year=isoyear(date),
         week=isoweek(date)) %>%
  select(CaseNumber, year, week, latitude, longitude) %>%
  distinct(CaseNumber, .keep_all = TRUE) %>%
  drop_na(latitude, longitude) %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = "NAD83", remove=F) %>%
  st_join(mpls_tract) %>% #spatial join neighborhoods
  st_drop_geometry() %>%
  drop_na(GEOID) %>%
  group_by(year, week, GEOID, .drop=F) %>%
  tally(name = "car_jack") %>%
  ungroup() %>%
  complete(year, week, GEOID=mpls_tract$GEOID, fill = list(car_jack = 0)) %>%
  filter(!(year==2021 & week==53)) %>% 
  arrange(GEOID, year, week) %>%
  left_join(mpls_tract, by = "GEOID") %>%
  left_join(pop_denom_locf, by = c("GEOID", "year")) %>%
  mutate(car_jack_rate = car_jack/total_pop*1000) %>%
  st_as_sf()
```

# MPLS Carjackings by Week - MPD Extended Data

```{r}
#aggregate to week over tracts
cj_exp_week <- cj_exp %>%
  group_by(year, week) %>%
  summarize(car_jack = sum(car_jack, na.rm = T),
            total_pop = sum(total_pop, na.rm = T)) %>%
  mutate(begin_date = ISOweek2date(paste(year, paste0("W", sprintf("%02d", week)), 1,sep = "-")),
         end_date = begin_date+weeks(1)-days(1),
         car_jack_rate = car_jack/total_pop*1000) 
  
ggplot(cj_exp_week)+
  geom_line(aes(x=begin_date, y=car_jack_rate))+ 
  scale_x_date(date_labels = "%b-%Y", 
               limits = c(min(cj_exp_week$begin_date), max(cj_exp_week$begin_date)))+
  geom_vline(xintercept=cj_exp_week$begin_date[cj_exp_week$year==2020 & 
                                                 cj_exp_week$week==isoweek(date("2020-09-22"))], 
              linetype="dotted", color="red", size=1)+
   geom_label(aes(x=cj_exp_week$begin_date[cj_exp_week$year==2020 &
                                             cj_exp_week$week==isoweek(date("2020-09-22"))],
                 y=0.050), 
             label = "Code Change", show.legend = FALSE)+
  labs(title = "Figure X: Weekly Minneapolis Carjackings, 1/1/2017-12/31/2021",
       x = "Week",
       y = "Weekly Carjacking Rate/ 1,000")+
  theme_minimal()+
    theme(axis.text.x=element_text(angle=45, hjust=1)) 
```


# MPLS ZCTA Carjackings Map - MPD Extended Data

```{r}
#aggregate to neighborhood-year level
cj_exp_tract_year <- cj_exp %>%
  group_by(GEOID, year) %>%
  summarize(car_jack = sum(car_jack, na.rm = T),
            total_pop = sum(B01001_001E, na.rm = T),
            car_jack_rate = car_jack/total_pop*1000) %>%
  mutate(GEOID = as.character(GEOID))


ggplot() +
  geom_sf(data = cj_exp_tract_year, aes(geometry = geometry, fill = car_jack_rate)) + 
  geom_sf(data = mpls, aes(geometry = geometry), color = "black", alpha = 0)+
  facet_grid(~year)+
  scale_fill_distiller(palette = "Spectral")+
  labs(title = "Figure X: Minneapolis Carjacking Rates by Tract and Year", 
       fill = "Weekly Rate/ 1,000")+
  theme_void()
```

#  MPLS Murder (Crime Incidents) Data

```{r}
#pre-pims
mpd_2016 <- read_csv("Data/Police_Incidents_2016.csv") 
mpd_2017 <- read_csv("Data/Police_Incidents_2017.csv") 
mpd_2018a <- read_csv("Data/Police_Incidents_2018.csv")

#pims
mpd_2018b <- read_csv("Data/Police_Incidents_2018_PIMS.csv") 
mpd_2019 <- read_csv("Data/Police_Incidents_2019.csv") 
mpd_2020 <- read_csv("Data/Police_Incidents_2020.csv") 
mpd_2021 <- read_csv("Data/Police_Incidents_2021.csv")

pre_pims_base <- mpd_2016 %>%
  rbind(mpd_2017) %>%
  rbind(mpd_2018a) %>%
  rename(reportedDate = ReportedDate,
         centerLong = Long,
         centerLat = Lat) %>%
  select(FID, centerLong, centerLat, Offense, reportedDate) %>%
  rename(OBJECTID = FID, 
         X = centerLong, 
         Y = centerLat, 
         offense = Offense)

post_pims_base <- mpd_2018b %>%
  rbind(mpd_2019) %>%
  rbind(mpd_2020) %>%
  rbind(mpd_2021) %>%
  select(OBJECTID, X, Y, offense, reportedDate)

mpd <- pre_pims_base %>%
  rbind(post_pims_base) 

#aggregate homicides to tract-week
homicide <- mpd %>%
  mutate(date=ymd_hms(reportedDate),
         year=isoyear(date),
         week=isoweek(date)) %>%
  filter(offense=="MURDR" & year!=2016 & year!=2015 & year!=2022) %>% #filter homicides
  select(OBJECTID, year, week, Y, X) %>%
  st_as_sf(coords = c("X", "Y"), crs = "NAD83", remove=F) %>%
  st_join(mpls_tract) %>% #spatial join neighborhoods
  st_drop_geometry() %>%
  filter(!is.na(GEOID)) %>%
  group_by(year, week, GEOID, .drop=F) %>%
  tally(name = "homicide") %>%
  ungroup() %>%
  complete(year, week, GEOID=mpls_tract$GEOID, fill = list(homicide = 0)) %>%
  filter(!(year==2021 & week==53)) %>% 
  arrange(GEOID, year, week) %>%
  left_join(mpls_tract, by = "GEOID") %>%
  left_join(pop_denom_locf, by = c("GEOID", "year")) %>%
  mutate(homicide_rate = homicide/total_pop*1000) %>%
  st_as_sf()
```

# MPLS Murder by Week 

```{r}
#aggregate to week over tracts
homicide_week <- homicide %>%
  group_by(year, week) %>%
  summarize(homicide = sum(homicide, na.rm = T),
            total_pop = sum(total_pop, na.rm = T)) %>%
  mutate(begin_date = ISOweek2date(paste(year, paste0("W", sprintf("%02d", week)), 1,sep = "-")),
         end_date = begin_date+weeks(1)-days(1),
         homicide_rate = homicide/total_pop*1000) 
  
ggplot(homicide_week)+
  geom_line(aes(x=begin_date, y=homicide_rate))+ 
  scale_x_date(date_labels = "%b-%Y", 
               limits = c(min(homicide_week$begin_date), max(homicide_week$begin_date)))+
     labs(title = "Figure X: Weekly Minneapolis Homicide, 1/1/2017-12/31/2021",
       x = "Week",
       y = "Weekly Homicide Rate/ 1,000")+
  theme_minimal()+
    theme(axis.text.x=element_text(angle=45, hjust=1)) 
```
# MPLS ZCTA Carjackings Map - MPD Extended Data

```{r}
#aggregate to neighborhood-year level
homicide_tract_year <- homicide %>%
  group_by(GEOID, year) %>%
  summarize(homicide = sum(homicide, na.rm = T),
            total_pop = sum(B01001_001E, na.rm = T),
            homicide_rate = homicide/total_pop*1000) %>%
  mutate(GEOID = as.character(GEOID))


ggplot() +
  geom_sf(data = homicide_tract_year, aes(geometry = geometry, fill = homicide_rate)) + 
  geom_sf(data = mpls, aes(geometry = geometry), color = "black", alpha = 0)+
  facet_grid(~year)+
  scale_fill_distiller(palette = "Spectral")+
  labs(title = "Figure X: Minneapolis Homicide Rates by Tract and Year", 
       fill = "Weekly Rate/ 1,000")+
  theme_void()
```
# Spatial Correlation of Carjackings and Homicide

```{r}
library(sfdep)

hom_2021 <- homicide_tract_year %>%
  filter(year==2021)

cj_2021 <- cj_exp_tract_year %>%
  filter(year==2021)

#hom_2021_sp <- as(hom_2021, "Spatial")

nb <- st_contiguity(hom_2021, queen=TRUE)
wt <-  st_weights(nb, style = "W")
global_moran(hom_2021$homicide_rate, nb, wt)
global_moran_test(
  hom_2021$homicide_rate,
  nb,
  wt,
  alternative = "greater",
  randomization = TRUE)

global_moran_test(
  cj_2021$car_jack_rate,
  nb,
  wt,
  alternative = "greater",
  randomization = TRUE)
```
# Spatial Heterogeneity in Car Jackings


```{r, message=F}
cj_exp_year_acs <- cj_exp_tract_year %>%
  left_join(acs_2020, by = c("GEOID")) %>%
  ungroup() %>%
  mutate(yearzero = year-2017,
         pov_rate_center = pov_rate-mean(pov_rate, na.rm = T),
         white_perc_center = scale(white_perc, center = T, scale = F),
         black_perc_center = scale(black_perc, center = T, scale = F),
         anyjack = ifelse(car_jack==0, 0, 1))

library(lme4)
library(lmerTest)

efa_vars <- cj_exp_year_acs %>%
  select(pov_rate, pub_assist_rate, female_hh_rate, 
         unemp_rate, age_below_18_perc, black_perc, 
         foreign_rate, hisp_perc, res_mob_rate, own_hh_rate) %>%
  st_drop_geometry()

fit <- factanal(efa_vars, factors = 3, rotation = "promax")
print(fit, digits = 2, cutoff = .6, sort = T)

library(lavaan)

cd_model_1  <- ' cd =~ unemp_rate + pov_rate +  female_hh_rate + no_hs_dip_rate + black_perc 
                  black_perc ~~ unemp_rate'  

cfa_cd <- cfa(cd_model_1, data = cj_exp_year_acs, std.lv = T)
modificationindices(cfa_cd)
summary(cfa_cd, fit.measures=TRUE, standardized = T)
cd_predict <- as.vector(lavPredict(cfa_cd, newdata = cj_exp_year_acs))
cj_exp_year_acs$conc_dis <- cd_predict


re <- lmer(car_jack_rate~as.factor(yearzero)*black_perc+
             (1|NAME), 
          data = cj_exp_year_acs)
summary(re)


re_logit <- glmer(anyjack ~ as.factor(yearzero)*black_perc + (1|NAME), 
                  data = cj_exp_year_acs, family = binomial)
summary(re_logit)


re_cd <- lmer(car_jack_rate~as.factor(yearzero)*conc_dis+
             (1|NAME), 
          data = cj_exp_year_acs)
summary(re)


re_logit_cd <- glmer(anyjack ~ as.factor(yearzero)*conc_dis + (1|NAME), 
                  data = cj_exp_year_acs, family = binomial)
summary(re_logit)
```

